Formatting Instructions for NIPS 2015

### 2015-11-09 writing/S'15 LSTM/data.tex
Camera ready for CoCo.

\begin{table}[tp]
  \centering\small
%  \begin{subtable}[t]{0.45\textwidth}
%    \centering
%    \begin{tabular}[t]{l l}
%      \toprule
%      Formula     & Interpretation \\
%      \midrule
%      $p_1$, $p_2$, $p_3$, $p_4$, $p_5$, $p_6$ & $\sem{x} \in \{\True, \False\}$ \\
%      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
%      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
%      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
%      \bottomrule
%    \end{tabular}    
%    \caption{Well-formed formulae. $\varphi$ and $\psi$
%      range over all well-formed formulae, and $\sem{\cdot}$ is
%      the interpretation function mapping formulae into $\{\True,
%      \False\}$.}\label{tab:pl}
%  \end{subtable}
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg p_3$        & $\natneg$ & $p_3$ \\
%      $\plneg \plneg p_6$ & $\nateq$  & $p_6$ \\
      $p_3$               & $\natfor$ & $p_3 \plor p_2$ \\
%      $p_1 \plor (p_2 \plor p_4)$               & $\natrev$ & $p_2 \pland  \plneg p_4$ \\
      $(\plneg p_2) \pland p_6  $ & $\natalt$ & $     \plneg ( p_6 \plor ( p_5 \plor p_3 ) ) $\\
      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
 $    p_4 \plor ( \plneg ( ( p_1  \plor p_6 )  \plor p_4 ) ) $ & $ \natfor $ & $  \plneg ( ( ( ( \plneg p_6 ) \plor ( \plneg p_4 ) )  \pland ( \plneg p_5 ) )\pland ( p_6 \pland p_6 ) ) $\\      
% $(p_3 \pland \plneg p_1 ) \plor \plneg p_3$    & $\natrev$& $\plneg\, (p_3 \plor p_2)$ \\
      %<	( not ( c ( or b ) ) )	( ( c ( and ( not a ) ) ) ( or ( not c ) ) )
      \bottomrule
    \end{tabular}
    \caption{Examples of short to moderate length pairs from the artificial data introduced in \protect\cite{Bowman:Potts:Manning:2014}. We only show the parentheses that are needed to disambiguate the sentences rather than the full binary bracketings.}\label{tab:plexs}
\end{table}

@@

\paragraph{The corpus}
The sentence pairs in the corpus are divided into thirteen bins according to the number of logical connectives (\word{and, or, not}) in the longer of the two sentences in the pair. We test the model on each bin separately (58k total examples, using an 80/20\% train/test split) in order to evaluate how its performance depends on the complexity of the sentences. In three experiments, we train our models on the training portions of bins 0--3 (62k examples), 0--4 (90k), and 0--6 (160k), and test on every bin but the trivial bin 0. Capping the size of the training sentences allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures. If their performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and L2 regularization (which helps overall performance) exacerbates this by discouraging the model from using the kind of numerically precise, nonlinearity-saturating functions that generalize best.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

>>>>

\begin{table}[tp]
  \centering\small
%  \begin{subtable}[t]{0.45\textwidth}
%    \centering
%    \begin{tabular}[t]{l l}
%      \toprule
%      Formula     & Interpretation \\
%      \midrule
%      $p_1$, $p_2$, $p_3$, $p_4$, $p_5$, $p_6$ & $\sem{x} \in \{\True, \False\}$ \\
%      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
%      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
%      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
%      \bottomrule
%    \end{tabular}    
%    \caption{Well-formed formulae. $\varphi$ and $\psi$
%      range over all well-formed formulae, and $\sem{\cdot}$ is
%      the interpretation function mapping formulae into $\{\True,
%      \False\}$.}\label{tab:pl}
%  \end{subtable}
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg p_3$        & $\natneg$ & $p_3$ \\
%      $\plneg \plneg p_6$ & $\nateq$  & $p_6$ \\
      $p_3$               & $\natfor$ & $p_3 \plor p_2$ \\
%      $p_1 \plor (p_2 \plor p_4)$               & $\natrev$ & $p_2 \pland  \plneg p_4$ \\
      $(\plneg p_2) \pland p_6  $ & $\natalt$ & $     \plneg ( p_6 \plor ( p_5 \plor p_3 ) ) $\\
      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
 $    p_4 \plor ( \plneg ( ( p_1  \plor p_6 )  \plor p_4 ) ) $ & $ \natfor $ & $  \plneg ( ( ( ( \plneg p_6 ) \plor ( \plneg p_4 ) )  \pland ( \plneg p_5 ) )\pland ( p_6 \pland p_6 ) ) $\\      
% $(p_3 \pland \plneg p_1 ) \plor \plneg p_3$    & $\natrev$& $\plneg\, (p_3 \plor p_2)$ \\
      %<	( not ( c ( or b ) ) )	( ( c ( and ( not a ) ) ) ( or ( not c ) ) )
      \bottomrule
    \end{tabular}
    \caption{Examples of short to moderate length pairs from the artificial data introduced in \protect\cite{Bowman:Potts:Manning:2014}. We only show the parentheses that are needed to disambiguate the sentences rather than the full binary bracketings that the models use.}\label{tab:plexs}
\end{table}

@@

\paragraph{The data}
Our sentence pairs are divided into thirteen bins according to the number of logical connectives (\word{and, or, not}) in the longer of the two sentences in each pair. We test each model on each bin separately (58k total examples, using an 80/20\% train/test split) in order to evaluate how each model's performance depends on the complexity of the sentences. In three experiments, we train our models on the training portions of bins 0--3 (62k examples), 0--4 (90k), and 0--6 (160k), and test on every bin but the trivial bin 0. Capping the size of the training sentences allows us to evaluate how the models interpret the sentences: if a model's performance falls off abruptly above the cutoff, it is reasonable to conclude that it relies heavily on specific sentence structures and cannot generalize to new structures. If a model's performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, all models will make errors on sentences above some length, and L2 regularization (which helps overall performance) exacerbates this by discouraging the model from using the kind of numerically precise, nonlinearity-saturating functions that generalize best.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

### 2015-11-09 writing/S'15 LSTM/discussion.tex
Camera ready for CoCo.

The results are shown in Figure~\ref{prop-results}. 
The tree models fit the training data well, reaching 98.9, 98.8, and 98.4\% overall accuracy respectively in the $\le$6 setting, with the LSTM underfitting slightly at 94.8\%. 
In that setting, all models generalized well to structures of familiar length, with the tree models all surpassing 97\% on examples of size 4, and the LSTM reaching 94.8\%.
On the longer test sentences, the tree models decayed smoothly in performance across the board, while the LSTM decayed more quickly and more abruptly, with a striking difference in the $\le$4 setting, where LSTM performance falls 10\% from 4 to 5, compared to 4.4\% for the next worse model. However, the LSTM improves considerably with more ample training data in the $\le$6 condition, showing only a 3\% drop and generalization results better than the best model's in the $\le$3 setting.

@@

We found that all four models were able to effectively exploit a recursively defined language to interpret sentences with complex unseen structures.
We find that tree models' biases allow them to do this with greater efficiency, outperforming sequence-based models substantially in every experiment. However, our sequence model was nonetheless able to generalize smoothly to from seen sentence structures to unseen ones, showing that its lack of explicit recursive structure did not prevent it from recognizing recursive structure in our artificial language.

>>>>

The results are shown in Figure~\ref{prop-results}. 
The tree models fit the training data well, reaching 98.9, 98.8, and 98.4\% overall accuracy respectively in the $\le$6 setting, with the LSTM underfitting slightly at 94.8\%. 
In that setting, all models generalized well to structures of familiar length, with the tree models all surpassing 97\% on examples in bin 4, and the LSTM reaching 94.8\%.
On the longer test sentences, the tree models decay smoothly in performance across the board, while the LSTM decays more quickly and more abruptly, with a striking difference in the $\le$4 setting, where LSTM performance falls 10\% from bin 4 to bin 5, compared to 4.4\% for the next worse model. However, the LSTM improves considerably with more ample training data in the $\le$6 condition, showing only a 3\% drop and generalization results better than the best model's in the $\le$3 setting.

@@

We find that all four models are able to effectively exploit a recursively defined language to interpret sentences with complex unseen structures.
We find that tree models' biases allow them to do this with greater efficiency, outperforming sequence-based models substantially in every experiment. However, our sequence model is nonetheless able to generalize smoothly from seen sentence structures to unseen ones, showing that its lack of explicit recursive structure does not prevent it from recognizing recursive structure in our artificial language.

### 2015-11-09 writing/S'15 LSTM/intro.tex
Camera ready for CoCo.

We propose a possible explanation for these results: standard sequence models can learn to
exploit recursive syntactic structure in generating sentence meaning representations, thereby 
learning to use the 
structure that tree models are explicitly designed around. This first requires that
sequence models are able to identify syntactic structure in natural language. We believe this is plausible, on the basis of other recent research \cite{vinyals2014grammar,Karpathy2015vaurn}.
In this paper, we  evaluate whether LSTM models are able to use that structure to guide interpretation, 
focusing on cases where the relevant syntactic structure is clearly indicated in the data.

We compare standard tree and sequence models on their handling of recursive structure by training the models on sentences whose length and recursion depth are limited, and then testing them on longer and more complex sentences, such that only models that exploit the recursive structure will be able to generalize in a way that yields correct interpretations for these test sentences. Our methods extend those of our earlier work in \cite{Bowman:Potts:Manning:2014}, which introduces an experiment and corresponding artificial dataset to test this ability in two tree models. We adapt that experiment to sequence models by decorating the statements with an explicit bracketing, and we use this design to compare an LSTM sequence model with three tree models, with a focus on what data each model needs in order to learn the needed generalizations.

@@

Our results engage with those of \cite{vinyals2014grammar} and \cite{dyer2015transition}, who find that sequence models can learn to recognize syntactic structure in natural language, at least when trained on explicitly syntactic tasks. The simplest model presented in \cite{vinyals2014grammar} uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) with high accuracy using only the information present in the vector. This shows that the LSTM is able to identify the correct syntactic structures and also hints that it is able to develop a generalizable method for encoding these structures in vectors. However, the massive size of the dataset needed to train the model---250M tokens---leaves open the possibility that it primarily learns to generate only tree structures that it has already seen, representing them as simple hashes---which would not capture unseen tree structures---rather than as structured objects.
Our experiments, though, show that LSTMs can learn to understand tree structures when given enough data, suggesting that there is no fundamental obstacle to learning this kind of structured representation. We also find, though, that sequence models lag behind tree models across the board, even on training corpora that are quite large relative to the complexity of the underlying grammar, suggesting that tree models can play a valuable role in tasks that require recursive interpretation.

>>>>

We propose a possible explanation for these results: standard sequence models can learn to
exploit recursive syntactic structure in generating representations of sentence meaning, thereby 
learning to use the 
structure that tree models are explicitly designed around. This requires that
sequence models be able to identify syntactic structure in natural language. We believe this is plausible on the basis of other recent research \cite{vinyals2014grammar,Karpathy2015vaurn}.
In this paper, we  evaluate whether LSTM sequence models are able to use such structure to guide interpretation, 
focusing on cases where syntactic structure is clearly indicated in the data.

We compare standard tree and sequence models on their handling of recursive structure by training the models on sentences whose length and recursion depth are limited, and then testing them on longer and more complex sentences, such that only models that exploit the recursive structure will be able to generalize in a way that yields correct interpretations for these test sentences. Our methods extend those of our earlier work in \cite{Bowman:Potts:Manning:2014}, which introduces an experiment and corresponding artificial dataset to test this ability in two tree models. We adapt that experiment to sequence models by decorating the statements with an explicit bracketing, and we use this design to compare an LSTM sequence model with three tree models, with a focus on what data each model needs in order to generalize well.

@@

Our results engage with those of \cite{vinyals2014grammar} and \cite{dyer2015transition}, who find that sequence models can learn to recognize syntactic structure in natural language, at least when trained on explicitly syntactic tasks. The simplest model presented in \cite{vinyals2014grammar} uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) with high accuracy using only the information present in the vector. This shows that the LSTM is able to identify the correct syntactic structures and also hints that it is able to develop a generalizable method for encoding these structures in vectors. However, the massive size of the dataset needed to train that model, 250M tokens, leaves open the possibility that it primarily learns to generate only tree structures that it has already seen, representing them as simple hashes---which would not capture unseen tree structures---rather than as structured objects.
Our experiments, though, show that LSTMs can learn to understand tree structures when given enough data, suggesting that there is no fundamental obstacle to learning this kind of structured representation. We also find, though, that sequence models lag behind tree models across the board, even on training corpora that are quite large relative to the complexity of the underlying grammar, suggesting that tree models can play a valuable role in tasks that require recursive interpretation.

### 2015-11-09 writing/S'15 LSTM/methods.tex
Camera ready for CoCo.

We use the architecture depicted in Figure~\ref{fig:model:top}, which builds on the one used in \newcite{Bowman:Potts:Manning:2014}. The model architecture uses two copies of a single sentence model (a tree or sequence model) to encode the premise and hypothesis (left and right side) expressions, and then uses those encodings as the features for a multilayer classifier which predicts one of the seven relations. Since the encodings are computed separately, the sentence models are forced to encode every element of the sentence meaning that the downstream model will need.

\begin{figure*}[t]
  \centering
  \input{figure1}
  \caption{In our model, two copies of a sentence model---based on either tree (b) or sequence (c) models---encode the two sentences. A multilayer classifier component (a) then uses the resulting vectors to predict a label that reflects the logical relationship between the two sentences.}
  \label{sample-figure}
\end{figure*}

@@

Our model is largely identical to the model from \newcite{Bowman:Potts:Manning:2014}, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all three models, rather than just the TreeRNTN model, so as to ensure that the three sentence models are compared in as similar a setting as possible.

@@

\paragraph{Sentence models}
The sentence encoding component of the model transforms the (jointly learned) embeddings of the input words into a single sentence vector. We experiment with tree-structured models (Figure~\ref{fig:model:tree}) with TreeRNN (eqn.~\ref{TreeRNN}), TreeRNTN (eqn.~\ref{TreeRNTN}), and TreeLSTM \cite{tai2015improved} activation functions. In addition, we use a sequence model (Figure~\ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} implemented as in \newcite{zaremba2015recurrent}. In experiments with a simpler non-LSTM RNN sequence model, the model tended to badly underfit the training data, and those results are not included here.

>>>>

We use the architecture depicted in Figure~\ref{fig:model:top}, which builds on the one used in \newcite{Bowman:Potts:Manning:2014}. The model architecture uses two copies of a single sentence model (a tree or sequence model) to encode the premise and hypothesis (left and right side) expressions, and then uses those encodings as the features for a multilayer classifier which predicts one of the seven relations. Since the encodings are computed separately, the sentence models must encode complete representations of the meanings of the two sentences for the downstream model to be able to succeed.

\begin{figure*}[t]
  \centering
  \input{figure1}
  \caption{In our model, two copies of a sentence model---based on either tree (b) or sequence (c) models---encode the two input sentences. A multilayer classifier component (a) then uses the resulting vectors to predict a label that reflects the logical relationship between the two sentences.}
  \label{sample-figure}
\end{figure*}

@@

Our model is largely identical to the model from \newcite{Bowman:Potts:Manning:2014}, but adds the two additional $\tanh$ NN layers, which we found help performance across the board, and also uses the NTN combination layer when evaluating all four models, rather than just the TreeRNTN model, so as to ensure that the sentence models are compared in as similar a setting as possible.

@@

\paragraph{Sentence models}
The sentence encoding component of the model transforms the (learned) embeddings of the input words for each sentence into a single vector representing that sentence. We experiment with tree-structured models (Figure~\ref{fig:model:tree}) with TreeRNN (eqn.~\ref{TreeRNN}), TreeRNTN (eqn.~\ref{TreeRNTN}), and TreeLSTM \cite{tai2015improved} activation functions. In addition, we use a sequence model (Figure~\ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} implemented as in \newcite{zaremba2015recurrent}. In experiments with a simpler non-LSTM RNN sequence model, the model tended to badly underfit the training data, and those results are not included here.
